Stack Safety as a Security Property

### 2020-07-04 main.tex
consistent use of Cref

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micropolicy enforcement
%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
Section~\ref{sec:relwork} discusses related work, and
Section~\ref{sec:future} sketches directions for future work.
%
We begin with a running example that we refer
to throughout the paper (Section~\ref{sec:running-example}) and some
technical preliminaries (Section~\ref{sec:prelim}).

@@

The program appears in Figure~\ref{fig:running-program}. It consists
of a {\tt main} function that performs a call to some function {\tt
  f}, which in turns makes a nested call to some function {\tt g}
before returning, all while performing some simple arithmetic
in the process.

@@

\bcp{We should also emphasize, again, that there is no heap (if there were,
we might put it above the stack or between the instructions and the stack)}
First of all, we ignore the heap. We are mainly concerned
with stack interactions so that will be the main focus. \leo{We're
  going to come back to that at the end. Will we?} We also
abstract away low-level details of memory layout and present a
machine that is word-addressed, with instructions starting from
location 0, a dedicated output at location 99, and lay the stack from
location 100 upwards. \leo{Explain more? How to connect to realistic
  one? Observations? DMA? Interrupt handlers?}
%
A sketch of the memory layout at that point can be found in the top of
Figure~\ref{fig:running-trace-b}: instructions for {\tt main} begin at
0, the dedicated output is empty, while the (uninitialized) stack
begins at 100. Directly below the layout, we summarize the parts of
the register file that are relevant to the example.
%
The rest of Figure~\ref{fig:running-trace-b} details the execution of the
program, showing the machine state at call and return points, as well
as describing the full effect of each instruction in detail.
\leo{Expand upon this?}

@@

\leo{TODO: Make attacker model more explicit} The execution trace
shown in Figure~\ref{fig:running-trace-b} is an example of a
``stack-safe'' execution: each process interacts only with the
registers, its local stack space and its call frame. But what could go
wrong if an attacker could hijack an arbitrary call and execute
arbitrary code? In the rest of this section we'll go over examples of
unsafe behaviors, before formally capturing what stack-safe means
formally and discussing enforcement techniques from the literature
in later sections.

@@

To set the stage, we first describe our machine model. Our definition
of stack safety is generic and this section discusses an abstract
interface to our machine model and some kind of static or dynamic,
software- or hardware-enforced security monitor.  For concreteness,
our running examples will use a conventional machine architecture and
a standard (RISC-V) instruction set; in~\Cref{sec:micropolicies} we
will extend this model it with a dynamic policy enforcement mechanism
based on \bcp{NickPaper,AndMaybeSomeOthers}.

@@

For an example see Figure \ref{fig:intex}, in which {\tt h} writes its return
value to {\tt g}'s return location the stack rather than its own. That address
is high integrity (red), and so writing to it causes the property to fail.

@@

Figure \ref{fig:confex2} shows the initial states \(\mach_e\), left,
and \(\mach_e'\), right. Low confidentiality data in green is shared
between them, while high confidentiality data in red and yellow may differ.
After executing the code shown, the final states below have written different
values to address 42, represented by the black and white components.
It is allowed for address 40 to differ between the states due to the initial
variation, but address 42 changed differently between the traces, representing
a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
%is read, it is overwritten before return, and the write at step 3 always writes
%the same value. Yellow and red locations differ between variations only because
%of the initial variation, and the rest are identical. So this example passes.

@@

Figure \ref{confex2} gives an example of a trace that fails trace
confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
Since {\tt x}, at address 100, is high confidentiality (red) the variant
may have a different value (yellow). So both \(r_4\) and address 107
differ in the final state between the original trace and the variant one,
failing the ending condition of trace confidentiality.

@@

      In Figure \ref{fig:lazyex1}, the final states of the original and variant
      differ, because a written value (in black or white) depended on the value
      of a high confidentiality location. So the resulting trace fails trace
      confidentiality. For observable trace confidentiality, we ask whether the
      differing values at address 42 cause any observable distinctions in
      future behavior. To do so, we must undo the variation of formerly high
      confidentiality locations, so that those in yellow match those in red,
      while preserving the updated values where data may have been leaked.

>>>>

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micropolicy enforcement
%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
\Cref{sec:relwork} discusses related work, and
\Cref{sec:future} sketches directions for future work.
%
We begin with a running example that we refer
to throughout the paper (\Cref{sec:running-example}) and some
technical preliminaries (\Cref{sec:prelim}).

@@

The program appears in \Cref{fig:running-program}. It consists
of a {\tt main} function that performs a call to some function {\tt
  f}, which in turns makes a nested call to some function {\tt g}
before returning, all while performing some simple arithmetic
in the process.

@@

\bcp{We should also emphasize, again, that there is no heap (if there were,
we might put it above the stack or between the instructions and the stack)}
First of all, we ignore the heap. We are mainly concerned
with stack interactions so that will be the main focus. \leo{We're
  going to come back to that at the end. Will we?} We also
abstract away low-level details of memory layout and present a
machine that is word-addressed, with instructions starting from
location 0, a dedicated output at location 99, and lay the stack from
location 100 upwards. \leo{Explain more? How to connect to realistic
  one? Observations? DMA? Interrupt handlers?}
%
A sketch of the memory layout at that point can be found in the top of
\Cref{fig:running-trace-b}: instructions for {\tt main} begin at
0, the dedicated output is empty, while the (uninitialized) stack
begins at 100. Directly below the layout, we summarize the parts of
the register file that are relevant to the example.
%
The rest of \Cref{fig:running-trace-b} details the execution of the
program, showing the machine state at call and return points, as well
as describing the full effect of each instruction in detail.
\leo{Expand upon this?}

@@

\leo{TODO: Make attacker model more explicit} The execution trace
shown in \Cref{fig:running-trace-b} is an example of a
``stack-safe'' execution: each process interacts only with the
registers, its local stack space and its call frame. But what could go
wrong if an attacker could hijack an arbitrary call and execute
arbitrary code? In the rest of this section we'll go over examples of
unsafe behaviors, before formally capturing what stack-safe means
formally and discussing enforcement techniques from the literature
in later sections.

@@

To set the stage, we first describe our machine model. Our definition
of stack safety is generic and this section discusses an abstract
interface to our machine model and some kind of static or dynamic,
software- or hardware-enforced security monitor.  For concreteness,
our running examples will use a conventional machine architecture and
a standard (RISC-V) instruction set; in \Cref{sec:micropolicies} we
will extend this model it with a dynamic policy enforcement mechanism
based on \bcp{NickPaper,AndMaybeSomeOthers}.

@@

For an example see \Cref{fig:intex}, in which {\tt h} writes its return
value to {\tt g}'s return location the stack rather than its own. That address
is high integrity (red), and so writing to it causes the property to fail.

@@

\Cref{fig:confex2} shows the initial states \(\mach_e\), left,
and \(\mach_e'\), right. Low confidentiality data in green is shared
between them, while high confidentiality data in red and yellow may differ.
After executing the code shown, the final states below have written different
values to address 42, represented by the black and white components.
It is allowed for address 40 to differ between the states due to the initial
variation, but address 42 changed differently between the traces, representing
a leak of data. %Meanwhile \ref{fig:lazyex2}, while high confidentiality data
%is read, it is overwritten before return, and the write at step 3 always writes
%the same value. Yellow and red locations differ between variations only because
%of the initial variation, and the rest are identical. So this example passes.

@@

\Cref{confex2} gives an example of a trace that fails trace
confidentiality. At instruction 42, {\tt h} reads {\tt x} from {\tt main}.
Since {\tt x}, at address 100, is high confidentiality (red) the variant
may have a different value (yellow). So both \(r_4\) and address 107
differ in the final state between the original trace and the variant one,
failing the ending condition of trace confidentiality.

@@

      In \Cref{fig:lazyex1}, the final states of the original and variant
      differ, because a written value (in black or white) depended on the value
      of a high confidentiality location. So the resulting trace fails trace
      confidentiality. For observable trace confidentiality, we ask whether the
      differing values at address 42 cause any observable distinctions in
      future behavior. To do so, we must undo the variation of formerly high
      confidentiality locations, so that those in yellow match those in red,
      while preserving the updated values where data may have been leaked.
